---
title: "Packages and data"
output: 
  pdf_document: 
    keep_tex: yes
---

This section is for opening packages and importing data.


```{r, setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, eval=FALSE, tidy.opts=list(width.cutoff=20), tidy=TRUE, wrap=TRUE)

```


The packages below are the ones used in the notebooks described in this project. 

```{r, message=FALSE}
library(tidyverse)
library(qiime2R)
library(metacoder)

```

Import your sample metadata.
```{r, message=FALSE}
metadata<-read_tsv("<your directory/sample-metadata.tsv>")
```

Import your taxonomy table, rooted tree, and feature table of choice.
```{r}
<xx_taxonomy> <-read_qza("<your directory/xx-taxonomy-rescript.qza>")$data
<xx_rooted_tree> <-read_qza("<your directory/xx-rooted-tree.qza>")$data
<xx_table> <-read_qza("<your directory/xx-table.qza>")$data
```

Import the relevant alpha diversity data for your needs. Name variable as needed and path as needed. The code creates the column "SampleID" from the row names in the data.
```{r,message=FALSE}
<xx_shannon> <-read_qza("<your directory/xx-core-metrics-results/shannon_vector.qza>")$data %>% rownames_to_column("SampleID")
```

Import the relevant Beta diversity data for your needs. Name variable as needed and path as needed. 
```{r}
<xx_wunifrac> <-read_qza("<your directory/xx-core-metrics-results/weighted_unifrac_pcoa_results.qza>")
```

